---
title: "Project1"
author: "Daniel Lassiter, Tom Muhlbaur, and Rachael Stryker"
date: "9/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generating Hypotheses:

Daniel's edit.

Daniel's second edit.

The variables that we decided to hone in on for our exploratory data analysis are:

-   Quantitative: CARS (\# cars w/hazmat), CARSDMG (\# hazmat cars that were damaged or derailed), CARSHZD (\# of cars that released hazmat), MONTH, DAY, TIMEHR, TIMEMIN, AMPM, TRNSPD, HIGHSPD

-   Qualitative: RR2 (second railroad involved), TYPE (accident type), TYPEQ (car type), Cause (manually assigned from CAUSE), STATION, WEATHER, VISIBLTY

### Loading libraries and data, settign directories, and processing data

```{r}
#Import libraries
library(ggplot2)
library(lattice)

#Set directories:
setwd('..')
wd <- getwd()

sourcedir <- paste0(wd, "/Source/")
traindir <-  paste0(wd, "/Data/TrainData")


# Source AccidentInput
setwd(sourcedir)
source("AccidentInput.R")
source("SPM_Panel.R")
source("PCAplots.R")

# Create a list of data frames for each year of accident data

acts <- file.inputl(traindir)

# Create a data frame with all accidents from all years from 2001 - 2019
# with columns that are consistent for all of these years

# Get a common set the variables

comvar <- intersect(colnames(acts[[1]]), colnames(acts[[8]]))

# the combined data frame

totacts <- combine.data(acts)

# Update the TYPE variable to contain more legible values

totacts$TYPE <- factor(totacts$TYPE, labels = c("Derailment", "HeadOn", "Rearend", "Side", "Raking", "BrokenTrain", "Hwy-Rail", "GradeX", "Obstruction", "Explosive", "Fire","Other","SeeNarrative" ))

# Update TYPEQ variable:
totacts$TYPEQ <- factor(totacts$TYPEQ, labels = c("Freight", "Passenger", "Commuter", "Work",  "Single", "CutofCars", "Yard", "Light", "Maint", "Other", "Other", "Other", "Other", "Other", "Other"))

# Update WEATHER variable:

totacts$WEATHER <- factor(totacts$WEATHER, labels = c("clear", "cloudy", "rain", "fog", "sleet", "snow"))

# Update the visibility variable

totacts$VISIBLTY <- factor(totacts$VISIBLTY, labels = c("dawn", "day", "dusk", "dark"))

# Update the cause variable to have more legible values


totacts$Cause <- rep(NA, nrow(totacts))

totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "M")] <- "(M) Miscellaneous Causes Not Otherwise Listed"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "T")] <- "(T) Rack, Roadbed and Structures"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "S")] <- "(S) Signal and Communication"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "H")] <- "(H) Train operation - Human Factors"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "E")] <- "(E) Mechanical and Electrical Failures"

totacts$Cause <- factor(totacts$Cause)

# Remove duplicates

totacts_DR <- totacts[!(duplicated(totacts[, c("INCDTNO", "YEAR", "MONTH", "DAY", "TIMEHR", "TIMEMIN")])),]


```

## Looking at casualties

### Create a data frame containing only accidents with one or more casualties. Use the variables "INCDTNO", "YEAR", "MONTH", "DAY", "TIMEHR", "TIMEMIN" to determine if there are duplicates in the accident reports with one or more casualties. Report number of duplicates.Show a box plot of casualities per accident per year.

```{r}
totacts_DR$Casualty <- totacts_DR$TOTINJ + totacts_DR$TOTKLD
totacts_wCasualties_DR <- subset.data.frame(totacts_DR, totacts$Casualty>0)


#Reset rownames (observation #s) for sequential numbering- otherwise they will remain the #s from totacts_wCasualties_DR
rownames(totacts_wCasualties_DR) <- NULL

plt <- ggplot(data = totacts_wCasualties_DR, aes(x = as.factor(YEAR), y = Casualty)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_grey(start = 0.5, end = 0.8) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Box Plots of Casualtiese") +
  labs(x = "Year", y = "Casualties")


ggplot_build(plt)$data

plt

```

```{r}
# ggpairs(totacts_wCasualties_DR[,c("cars", "EQPDMG", "ACCDMG", "TOTINJ", "TOTKLD")])

```

### Bar plots of total # of accidents with 1 or more casualties for all quantitative varaibles.


```{r}
library(ggplot2)

# Looking at accidents with 1 or more casualty
ylab <- "Accidents with 1 or more casualty"

```

```{r}
## Accident type
tbl <- as.data.frame(table(totacts_wCasualties_DR$TYPE))
tbl
xlab <- "Type"
ggplot(as.data.frame(tbl), aes(x = Var1, y= Freq)) + xlab(xlab) + ylab(ylab)+ geom_bar(stat="identity")
```
* Highway-rail and derailments are the two accident types that most frequently yield one or more casualties.

```{r}
## Cause
tbl <- as.data.frame(table(totacts_wCasualties_DR$Cause))
xlab <- "Cause"
ggplot(as.data.frame(tbl), aes(x = Var1, y= Freq)) + xlab(xlab) + ylab(ylab)+ geom_bar(stat="identity")
```
* Miscellaneous causes was by far the most frequent accident type followed by human factors.

```{r}
## TypeQ
tbl <- as.data.frame(table(totacts_wCasualties_DR$TYPEQ))
xlab <- "TYPEQ"
ggplot(as.data.frame(tbl), aes(x = Var1, y= Freq)) + xlab(xlab) + ylab(ylab)+ geom_bar(stat="identity")
```
* Passenger cars and commuter cars for accidents with casualties.

```{r}
## Station
tbl <- as.data.frame(table(totacts_wCasualties_DR$STATION))
xlab <- "STATION"
ggplot(as.data.frame(tbl), aes(x = Var1, y= Freq)) + xlab(xlab) + ylab(ylab)+ geom_bar(stat="identity")
```
* No single station has more than 50 accidents with casualties so there's probably not enough data to build a model of casualties using station as a varaible.


```{r}
## WEATHER
tbl <- as.data.frame(table(totacts_wCasualties_DR$WEATHER))
xlab <- "WEATHER"
ggplot(as.data.frame(tbl), aes(x = Var1, y= Freq)) + xlab(xlab) + ylab(ylab)+ geom_bar(stat="identity")
```
* By far the majority of accidents occurred during clear weather.


```{r}
## VISIBLTY
tbl <- as.data.frame(table(totacts_wCasualties_DR$VISIBLTY))
xlab <- "VISIBLTY"
ggplot(as.data.frame(tbl), aes(x = Var1, y= Freq)) + xlab(xlab) + ylab(ylab)+ geom_bar(stat="identity")
```
* Most accidents occurred during the day.





### Zooming in on highway-rail type accidents: Interaction Plots between quantitative and qualitative variables.

```{r}

library(MASS)
library(lindia)
library(ggplot2)

# Create Night/Not Night as a variable
dark <- rep(0, nrow(totacts_wCasualties_DR))
dark[which(totacts_wCasualties_DR$VISIBLTY =='dark')] <- 1
dark <- as.factor(dark)
contrasts(dark)

# Create highway-rail variable
table(totacts_wCasualties_DR$TYPE)

HwyRail <- rep(0, nrow(totacts_wCasualties_DR))
HwyRail[which(totacts_wCasualties_DR$TYPE == 'Hwy-Rail')] <- 1 
HwyRail <- as.factor(HwyRail)
contrasts(HwyRail)

interaction.plot(HwyRail, dark, totacts_wCasualties_DR$Casualty)

# Interaction between trainspeed and highway-rail accidents
TrnSpd_box <- ggplot(xdmg, aes(y=TRNSPD)) + geom_boxplot()

med <- ggplot_build(TrnSpd_box)$data[[1]]$middle

#Create TRNSPD variable
TRNSPD.factor <- xdmg$TRNSPD
TRNSPD.factor[which(xdmg$TRNSPD<med)]<-'low train speed'
TRNSPD.factor[which(xdmg$TRNSPD>=med)]<-'high train speed'
TRNSPD.factor <- factor(TRNSPD.factor)
contrasts(TRNSPD.factor)

# Create CARS variable
CARS <- rep(0, nrow(xdmg))
CARS[which(xdmg$CARS > 0)] <- 1
CARS <- as.factor(CARS)
contrasts(CARS)

# Create Freight as variable
Freight <- rep(0, nrow(xdmg))
Freight[which(xdmg$TYPEQ =='Freight')] <- 1
Freight <- as.factor(Freight)
contrasts(Freight)

# Create derailment variable
Derail <- rep(0, nrow(xdmg))
Derail[which(xdmg$TYPE == 'Derailment')] <- 1 
Derail <- as.factor(Derail)
contrasts(xdmg$Derail)

# create an interaction plot for TRNSPD and Derailments
interaction.plot(TRNSPD.factor, CARS, log(xdmg$ACCDMG))
interaction.plot(TRNSPD.factor, Freight, log(xdmg$ACCDMG))
interaction.plot(TRNSPD.factor, Derail, log(xdmg$ACCDMG))

```


## Looking at extreme accidents

```{r}
dmgbox <- ggplot(totacts_DR, aes(y=ACCDMG)) + geom_boxplot()
dmgbox

# Names associated with box plot features:
names(ggplot_build(dmgbox)$data[[1]])

# ymax is the upper whisker - anything above that is an outlier
upper <- ggplot_build(dmgbox)$data[[1]]$ymax

# create a new data frame with only the outliers
xdmg <- totacts_DR[totacts_DR$ACCDMG > upper,]

# how many outliers are there
nrow(xdmg)

# What proportion of accidents are extreme?

frac_acts_x <- round(nrow(xdmg)/nrow(totacts), 2)*100

# Proportion of costs

frac_cost_x <- round(sum(as.numeric(totacts$ACCDMG[which(totacts$ACCDMG > ggplot_build(dmgbox)$data[[1]]$ymax)]))/sum(as.numeric(totacts$ACCDMG)), 2)*100

# Remove 9/11 accident



```

There are `r nrow(xdmg)` outliers which comprise `r frac_cost_x`% of the sum of accident damage across all accidents. `r frac_acts_x`% of all accidents are extreme accidents.

### Make a barplot of the cause of accidents for extreme accidents.

```{r}
# Use table() to see the frequencies

tbl <- as.data.frame(table(xdmg$Cause))

# Use barplot() to graph this

ggplot(tbl, aes(x = Var1, y= Freq)) + geom_bar(stat="identity")

tbl
```

### Make a barplot of the cause of accidents for extreme accidents.

```{r}
# Use table() to see the frequencies

tbl <- as.data.frame(table(xdmg$TYPE))

# Use barplot() to graph this

ggplot(tbl, aes(x = Var1, y= Freq)) + geom_bar(stat="identity")

tbl
```

### Box plot of extreme accidents by accident type

# ACCDMG Analysis:

# Casualties Analysis:
