---
title: "Project1"
author: "Daniel Lassiter, Tom Muhlbaur, and Rachael Stryker"
date: "9/28/2020"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Generating Hypotheses:
The variables that we decided to hone in on for our exploratory data analysis are:

* Quantitative: CARS (# cars w/hazmat), CARSDMG (# hazmat cars that were damaged or derailed), CARSHZD (# of cars that released hazmat), MONTH, DAY, TIMEHR, TIMEMIN, AMPM, TRNSPD, HIGHSPD
  
* Qualitative: RR2 (second railroad involved), TYPE (accident type), TYPEQ (car type), Cause (manually assigned from CAUSE), STATION, WEATHER, VISIBLTY
Extreme Damage:

Hypothesis 1: 
Hypothesis 2:

Casualties:

Hypothesis 1: The Chicago Station increases the number of casualties. (Null): The Chicago station does not increase the number of casualties
Hypothesis 2:

### Loading libraries and data, settign directories, and processing data

```{r}
#Import libraries
library(ggplot2)
library(lattice)
library(plyr)

#Set directories:
setwd('..')
wd <- getwd()

sourcedir <- paste0(wd, "/Source/")
traindir <-  paste0(wd, "/Data/TrainData")


# Source AccidentInput
setwd(sourcedir)
source("AccidentInput.R")
source("SPM_Panel.R")
source("PCAplots.R")

# Create a list of data frames for each year of accident data

acts <- file.inputl(traindir)

# Create a data frame with all accidents from all years from 2001 - 2019
# with columns that are consistent for all of these years

# Get a common set the variables

comvar <- intersect(colnames(acts[[1]]), colnames(acts[[8]]))

# the combined data frame

totacts <- combine.data(acts)

# Update the TYPE variable to contain more legible values

totacts$TYPE <- factor(totacts$TYPE, labels = c("Derailment", "HeadOn", "Rearend", "Side", "Raking", "BrokenTrain", "Hwy-Rail", "GradeX", "Obstruction", "Explosive", "Fire","Other","SeeNarrative" ))

# Update WEATHER variable:

totacts$WEATHER <- factor(totacts$WEATHER, labels = c("clear", "cloudy", "rain", "fog", "sleet", "snow"))

# Update the visibility variable

totacts$VISIBLTY <- factor(totacts$VISIBLTY, labels = c("dawn", "day", "dusk", "dark"))

# Update the cause variable to have more legible values

# Accident cause
totacts$Cause <- rep(NA, nrow(totacts))

totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "M")] <- "(M) Miscellaneous Causes Not Otherwise Listed"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "T")] <- "(T) Rack, Roadbed and Structures"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "S")] <- "(S) Signal and Communication"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "H")] <- "(H) Train operation - Human Factors"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "E")] <- "(E) Mechanical and Electrical Failures"

# This new variable, Cause, has to be a factor
totacts$Cause <- factor(totacts$Cause)

```
## Looking at casualties
### Create a data frame containing only accidents with one or more casualties. Use the variables "INCDTNO", "YEAR", "MONTH", "DAY", "TIMEHR", "TIMEMIN" to determine if there are duplicates in the accident reports with one or more casualties. Report number of duplicates.Show a box plot of casualities per accident per year.

```{r}
totacts$Casualty <- totacts$TOTINJ + totacts$TOTKLD
totacts_wCasualties <- subset.data.frame(totacts, totacts$Casualty>0)

# Remove 9/11
totacts_wCasualties911 <- totacts_wCasualties[-186,]

totacts_wCasualties_DR <- totacts_wCasualties[!(duplicated(totacts_wCasualties911[, c("INCDTNO", "YEAR", "MONTH", "DAY", "TIMEHR", "TIMEMIN")])),]

#Reset rownames (observation #s) for sequential numbering- otherwise they will remain the #s from totacts_wCasualties_DR
rownames(totacts_wCasualties_DR) <- NULL

plt <- ggplot(data = totacts_wCasualties_DR, aes(x = as.factor(YEAR), y = Casualty)) +
  geom_boxplot() +
  coord_flip() +
  scale_fill_grey(start = 0.5, end = 0.8) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Box Plots of Casualtiese") +
  labs(x = "Year", y = "Casualties")


ggplot_build(plt)$data

plt

```




```{r}
# ggpairs(totacts_wCasualties_DR[,c("cars", "EQPDMG", "ACCDMG", "TOTINJ", "TOTKLD")])

```



### Make a barplot of the type of accidents for all accidents with one or more casualties.
```{r}
library(ggplot2)

# Use table() to see the frequencies

table(totacts_wCasualties_DR$TYPE)

# Use barplot() to graph this

ggplot(as.data.frame(table(totacts_wCasualties_DR$TYPE)), aes(x = Var1, y= Freq)) + geom_bar(stat="identity")

```

Highway-rail and derailments are the two accident types that most frequently yield one or more casualties.  

### Make a barplot of the cause of accidents for all accidents with one or more casualties.
```{r}
# Use table() to see the frequencies

tbl <- as.data.frame(table(totacts_wCasualties_DR$Cause))

# Use barplot() to graph this

ggplot(tbl, aes(x = Var1, y= Freq)) + geom_bar(stat="identity")

tbl
```

Miscellaneous causes was by far the most frequent accident type  followed by human factors.

## Next, we will look at how the Station closest to the accident contributes to the number of events that have casualties.
```{r}
# Use count to see the frequency of occurrences of casualties at each Station

y = count(totacts_wCasualties_DR, "STATION")

statcas <- aggregate(cbind(Count, Casualty) ~ STATION, transform(totacts_wCasualties_DR, Count = 1), sum)
statcas

#Make a New variable Ratio
statcas$Ratio <- rep(NA, nrow(statcas))
statcas <- transform(statcas, Ratio = Casualty / Count)
# This new variable, Cause, has to be a factor

statcas$Ratio <- factor(statcas$Ratio)


# Use table() to look at the Stations with more than nine occurrences
xstat10 <- statcas[statcas$Count > 9,]
#table(xstat10)
#xstat10
class(xstat10)

# Use ggplot() to plot the frequency of casualties at the etreme stations./
ggplot(xstat10, aes(x = STATION, y= freq)) + geom_bar(stat="identity") + theme(text = element_text(size = 5.5))
ggplot(statcas, aes(x = STATION, y= Ratio)) + geom_bar(stat="identity") + theme(text = element_text(size = 5.5))
ggplot(xstat10, aes(x = STATION, y= Ratio)) + geom_bar(stat="identity") + theme(text = element_text(size = 5.5))
```

```{r}
#boxplots of casualties conditioned on the stations who had a frequency of 10 or more accidents that resulted in a casualty. 
bwplot(xstat10$STATION~ (totacts_wCasualties_DR$Casualty), main = "Box Plots of Stations with Frequency of Casualties", xlab= "Casualty" )
xstat10$STATION

```

```{r}
#Total number of frequencies of casualties at a station relative to the total of other stations' occurences.
xstatchic <- y[y$STATION == "CHICAGO", ]
sumfreq <- sum(y$freq)
totoccur <- xstatchic$freq / sumfreq
totoccur
#Only makes up for 1.3% of total occurrences, but that is out of 3417 different stations. Let's look at it compared to with stations with 5 or more occurrences. 
xstat1 <- y[y$freq > 4,]
sumfreq1 <- sum(xstat1$freq)
totoccur1 <- xstatchic$freq / sumfreq1
totoccur1
# Makes up for about six percent of occurrences out of 662 stations who have 5 or more occurences of casualties.
```

## Looking at extreme accidents

```{r}
dmgbox <- ggplot(totacts, aes(y=ACCDMG)) + geom_boxplot()
dmgbox

# Names associated with box plot features:
names(ggplot_build(dmgbox)$data[[1]])

# ymax is the upper whisker - anything above that is an outlier
upper <- ggplot_build(dmgbox)$data[[1]]$ymax

# create a new data frame with only the outliers
xdmg <- totacts[totacts$ACCDMG > upper,]

# how many outliers are there
nrow(xdmg)

# What proportion of accidents are extreme?

frac_acts_x <- round(nrow(xdmg)/nrow(totacts), 2)*100

# Proportion of costs

frac_cost_x <- round(sum(as.numeric(totacts$ACCDMG[which(totacts$ACCDMG > ggplot_build(dmgbox)$data[[1]]$ymax)]))/sum(as.numeric(totacts$ACCDMG)), 2)*100

```

There are `r nrow(xdmg)` outliers which comprise `r frac_cost_x`% of the sum of accident damage across all accidents. `r frac_acts_x`% of all accidents are extreme accidents.


### Make a barplot of the cause of accidents for extreme accidents.
```{r}
# Use table() to see the frequencies

tbl <- as.data.frame(table(xdmg$Cause))

# Use barplot() to graph this

ggplot(tbl, aes(x = Var1, y= Freq)) + geom_bar(stat="identity")

tbl
```


### Make a barplot of the cause of accidents for extreme accidents.
```{r}
# Use table() to see the frequencies

tbl <- as.data.frame(table(xdmg$TYPE))

# Use barplot() to graph this

ggplot(tbl, aes(x = Var1, y= Freq)) + geom_bar(stat="identity")

tbl
```

### Box plot of extreme accidents by accident type




# ACCDMG Analysis:


# Casualties Analysis:


