---
title: "Project1"
author: "Daniel Lassiter, Tom Muhlbaur, and Rachael Stryker"
date: "9/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generating Hypotheses:

The variables that we decided to hone in on for our exploratory data analysis are:

-   Quantitative: CARS (\# cars w/hazmat), CARSDMG (\# hazmat cars that were damaged or derailed), CARSHZD (\# of cars that released hazmat), MONTH, DAY, TIMEHR, TIMEMIN, AMPM, TRNSPD, HIGHSPD

-   Qualitative: RR2 (second railroad involved), TYPE (accident type), TYPEQ (car type), Cause (manually assigned from CAUSE), STATION, WEATHER, VISIBLTY

# Loading libraries and data, settign directories, and processing data

```{r}
#Import libraries
library(tidyverse)
library(tibble)
library(dplyr)
library(ggplot2)
library(ggfortify)
library(lattice)
library(GGally)
library(MASS)
library(lindia)
library(ggpubr)
# install.packages("devtools")
library(devtools)
# install_github("vqv/ggbiplot")
library(ggbiplot)
library(data.table)
library(gplots)
library(boot)
source("http://www.phaget4.org/R/myImagePlot.R")

#Set directories:
setwd('..')
wd <- getwd()

sourcedir <- paste0(wd, "/Source/")
traindir <-  paste0(wd, "/Data/TrainData")


# Source AccidentInput
setwd(sourcedir)
source("AccidentInput.R")
source("SPM_Panel.R")
source("PCAplots.R")

# Data Procesing

## Create list of dataframes
acts <- file.inputl(traindir)


## Combine all data into a single dataframe

totacts <- combine.data(acts)

## Process data variables to be easier to work with

### TYPE
totacts$TYPE <- factor(totacts$TYPE, labels = c("Derailment", "HeadOn", "Rearend", "Side", "Raking", "BrokenTrain", "Hwy-Rail", "GradeX", "Obstruction", "Explosive", "Fire","Other","SeeNarrative" ))

### TYPEQ
totacts$TYPEQ <- factor(totacts$TYPEQ, labels = c("Freight", "Passenger", "Commuter", "Work",  "Single", "CutofCars", "Yard", "Light", "Maint", "Other", "Other", "Other", "Other", "Other", "Other"))

### WEATHER
totacts$WEATHER <- factor(totacts$WEATHER, labels = c("clear", "cloudy", "rain", "fog", "sleet", "snow"))

### VISIBLTY
totacts$VISIBLTY <- factor(totacts$VISIBLTY, labels = c("dawn", "day", "dusk", "dark"))

### Cause
totacts$Cause <- rep(NA, nrow(totacts))

totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "M")] <- "(M) Miscellaneous Causes Not Otherwise Listed"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "T")] <- "(T) Rack, Roadbed and Structures"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "S")] <- "(S) Signal and Communication"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "H")] <- "(H) Train operation - Human Factors"
totacts$Cause[which(substr(totacts$CAUSE, 1, 1) == "E")] <- "(E) Mechanical and Electrical Failures"

totacts$Cause <- factor(totacts$Cause)

### Multi Railroad
totacts$MultiRR <- rep(NA, nrow(totacts))
totacts$MultiRR[which(totacts$RR2 == "")] <- FALSE
totacts$MultiRR[which(totacts$RR2 != "")] <- TRUE

### Time of day

for (i in 1:length(totacts)){
  if (totacts$AMPM[i] == 'PM'){
    totacts$Time24hr[i] <- totacts$TIMEHR[i] + 12 + totacts$TIMEMIN[i]/60
  }
  else{
    totacts$Time24hr[i] <- totacts$TIMEHR[i] + totacts$TIMEMIN[i]/60
  }
}

## Remove duplicates
totacts_DR <- totacts[!(duplicated(totacts[, c("INCDTNO", "YEAR", "MONTH", "DAY", "TIMEHR", "TIMEMIN")])),]

## Create dataframe with only accidents with 1 or more casualty
totacts_DR$Casualty <- (totacts_DR$TOTINJ + totacts_DR$TOTKLD)
totacts_wCasualties_DR <- subset.data.frame(totacts_DR, totacts_DR$Casualty>0)
rownames(totacts_wCasualties_DR) <- NULL

## Creating extreme accident cost dataframe
plt_box <- ggplot(totacts_DR, aes(y=ACCDMG)) + geom_boxplot()
upper <- ggplot_build(plt_box)$data[[1]]$ymax
xdmg <- totacts_DR[totacts_DR$ACCDMG > upper,]

### Remove 9/11
xdmg <- xdmg[-186, ]

## Creating extreme casualties dataframe
plt_box <- ggplot(totacts_wCasualties_DR, aes(y=Casualty)) + geom_boxplot()
upper <- ggplot_build(plt_box)$data[[1]]$ymax
xcas <- totacts_DR[totacts_DR$ACCDMG > upper,]
```

# Exploratory Data Analysis for Accidents with Casualties

```{r}
# Exploration:
## Accident frequency by type
ggplot(as.data.frame(table(totacts_wCasualties_DR$TYPE)), aes(x = Var1, y= Freq)) + geom_bar(stat="identity")
print("Most accidents with 1 or more casualties are of type highway-rail")

## Total casualties by type:
ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty)) + geom_col(na.rm=TRUE)
print("Highway-rail is also the category with the most number of casualties, but when comparing this plot with the previous, derailments seem to be more severe per accident.")

## Inspecting extreme events
plt_box <- ggplot(totacts_wCasualties_DR, aes(y=Casualty)) + geom_boxplot()
plt_box
upper <- ggplot_build(plt_box)$data[[1]]$ymax
print("Upper whisker:")
upper

tib <- as_tibble(totacts_wCasualties_DR)

print("Number of extreme accidents with more than 3 casualties")
count(tib %>% filter(Casualty > upper))

print("Fraction of total accidents that are above upper whisker:")
# count(tib %>% filter(Casualty > upper))/count(tib)

print("Fraction of total casualties that occurred in extreme accidents:")
tot_cas <- tib %>% summarize(Casualties = sum(Casualty))
xtrm_cas <- tib %>% filter(Casualty > upper) %>% summarize(Casualties = sum(Casualty))
xtrm_cas/tot_cas
print("This means that 10% of the accidents are responsible for 64% of casualties.")


## Inspecting for potential relationships between categorical variables
totacts_wCasualties_DR$Cas_factor <- rep(NA, nrow(totacts_wCasualties_DR))

totacts_wCasualties_DR$Cas_factor[which(totacts_wCasualties_DR$Casualty <= 3)] <- "Less than or equal to 3 Casualties"
totacts_wCasualties_DR$Cas_factor[which(totacts_wCasualties_DR$Casualty > 3)] <- "Greater  to 3 Casualties (Above upper whisker)"
totacts_wCasualties_DR$Cas_factor <- factor(totacts_wCasualties_DR$Cas_factor)

### (SUM) What other conditions are present for each casualty?
plt_1 <- ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty,fill=Cause)) + geom_col(na.rm=TRUE)

plt_2 <- ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty,fill=WEATHER)) + geom_col(na.rm=TRUE)

plt_3 <- ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty,fill=VISIBLTY)) + geom_col(na.rm=TRUE)

plt_4 <- ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty,fill=Cas_factor)) + geom_col(na.rm=TRUE)

plt_5 <- ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty,fill=TYPEQ)) + geom_col(na.rm=TRUE)

plt_6 <- ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty,fill=MultiRR)) + geom_col(na.rm=TRUE)
### (COUNT) What other conditions are present for each accident with a casualty?
plt_7 <- ggplot(totacts_wCasualties_DR,aes(TYPE,fill=Cause)) + geom_bar(na.rm=TRUE)

plt_8 <- ggplot(totacts_wCasualties_DR,aes(TYPE,fill=WEATHER)) + geom_bar(na.rm=TRUE)

plt_9 <- ggplot(totacts_wCasualties_DR,aes(TYPE,fill=VISIBLTY)) + geom_bar(na.rm=TRUE)

plt_10 <- ggplot(totacts_wCasualties_DR,aes(TYPE,fill=Cas_factor)) + geom_bar(na.rm=TRUE)

plt_11 <- ggplot(totacts_wCasualties_DR,aes(TYPE,fill=TYPEQ)) + geom_bar(na.rm=TRUE)

plt_12 <- ggplot(totacts_wCasualties_DR,aes(TYPE,Casualty,fill=MultiRR)) + geom_col(na.rm=TRUE)

ggarrange(plt_1, plt_2, plt_3, plt_4, plt_5, plt_6, nrow = 2, ncol = 3)
ggarrange(plt_7, plt_8, plt_9, plt_10, plt_11, plt_12, nrow = 2, ncol = 3)

print("There is a possible interaction between Derailments and Dark Visibility vs. Casualties and Derailments and Cause - Track, Roadbed, and Structures vs. Casualties. Also, unsurprisingly, most accidents with casualties and most total casualties occur on passenger and commuter trains primarily.")


## Looking for correlation between quantitative variables and casualties:
### Variables tested: "Casualty", "TRNSPD", "TONS", "CARS", "TEMP", "POSITON2", "HEADEND1", "HEADEND2", "MIDMAN1", "MIDMAN2", "LOADF1", "LOADF2", "LOADP1", "LOADP2", "ENGRS", "FIREMEN", "CONDUCTR", "BRAKEMEN", "Latitude", "Longitud"

uva.pairs(totacts_wCasualties_DR[, c("Casualty", "TRNSPD", "TEMP", "TONS", "LOADP1", "LOADP2", "CARS", "CARSHZD", "Time24hr")])
print("These are the four variables with the highest correlations coefficient. Unsurprisingly, the LOADP2 variable, number of derailed loaded passengar cars, has the highest correlation with casualty. This is another indicator that type derailment accidents are the most severe and warrant the closest scrutiny.")


## Doing PCA to look for hidden quantitative relationships:
# Perform PCA using the correlation matrix
totacts_wCasualties.pca <- princomp(totacts_wCasualties_DR[,c("Casualty", "TRNSPD", "CARS", "Time24hr", "TEMP", "TONS", "MONTH")], cor = T)

ggbiplot(totacts_wCasualties.pca, varname.size = 5, labels=row(totacts_wCasualties_DR)[,1], main="Biplot")

### Inspecting outliers:
totacts_wCasualties_DR[191, ]


### Removing outliers:

df <- totacts_wCasualties_DR[-191,]

totacts_wCasualties.pca <- princomp(df[,c("Casualty", "TRNSPD", "CARS", "Time24hr", "TEMP", "TONS", "MONTH")], cor = T)

ggbiplot(totacts_wCasualties.pca, varname.size = 5, labels=NULL[,1], main="Biplot", alpha = 0.1)


### Loadings on first 2 PCs
barplot(totacts_wCasualties.pca$loadings[,1], main="PC1 Loadings")
barplot(totacts_wCasualties.pca$loadings[,2], main='PC2 Loadings')
barplot(totacts_wCasualties.pca$loadings[,3], main='PC3 Loadings')

print("The loading on casualty is relatively insignificant until the 3rd PC where it moves with trainspeed, # of hazard cars, and 24 hour time of day.")


## Interaction plots
### Decided to work with the dataframe with the most extreme accident removed
totacts_wCasualties_DR_ER <- df



### Define qualitative variables
  #### Create Day/Not Day as a variable
day <- rep(0, nrow(totacts_wCasualties_DR_ER))
day[which(totacts_wCasualties_DR_ER$VISIBLTY =='day')] <- 1
day <- as.factor(day)
contrasts(day)

  #### Clear/not clear as variable
clear <- rep(0, nrow(totacts_wCasualties_DR_ER))
clear[which(totacts_wCasualties_DR_ER$WEATHER =='clear')] <- 1
clear <- as.factor(clear)
contrasts(clear)

  #### Create highway-rail variable
HwyRail <- rep(0, nrow(totacts_wCasualties_DR_ER))
HwyRail[which(totacts_wCasualties_DR_ER$TYPE == 'Hwy-Rail')] <- 1 
HwyRail <- as.factor(HwyRail)
contrasts(HwyRail)

  #### Create derailment variable
Derail <- rep(0, nrow(totacts_wCasualties_DR_ER))
Derail[which(totacts_wCasualties_DR_ER$TYPE == 'Derailment')] <- 1 
Derail <- as.factor(Derail)
contrasts(Derail)

  #### Create passenger or commuter variable
Pass_or_Comm <- rep(0, nrow(totacts_wCasualties_DR_ER))
Pass_or_Comm[which(totacts_wCasualties_DR_ER$TYPEQ == 'Passenger' | totacts_wCasualties_DR_ER$TYPEQ == 'Commuter' )] <- 1 
Pass_or_Comm <- as.factor(Pass_or_Comm)
contrasts(Pass_or_Comm)

  #### Cause - Train roadbed and structures
clear <- rep(0, nrow(totacts_wCasualties_DR_ER))
clear[which(totacts_wCasualties_DR_ER$WEATHER =='clear')] <- 1
clear <- as.factor(clear)
contrasts(clear)


### Quantitative Variables

  #### Create TRNSPD variable
TrnSpd_box <- ggplot(totacts_wCasualties_DR_ER, aes(y=TRNSPD)) + geom_boxplot()

med <- ggplot_build(TrnSpd_box)$data[[1]]$middle

TRNSPD.factor <- totacts_wCasualties_DR_ER$TRNSPD
TRNSPD.factor[which(totacts_wCasualties_DR_ER$TRNSPD<med)]<-'low train speed'
TRNSPD.factor[which(totacts_wCasualties_DR_ER$TRNSPD>=med)]<-'high train speed'
TRNSPD.factor <- factor(TRNSPD.factor)
contrasts(TRNSPD.factor)

  #### Create CARS variable
CARS.factor <- totacts_wCasualties_DR_ER$CARS
CARS.factor[which(totacts_wCasualties_DR_ER$CARS == 0)]<-'No Hazard Cars'
CARS.factor[which(totacts_wCasualties_DR_ER$CARS > 0)]<-'1 or More Hazard Cars'
CARS.factor <- factor(CARS.factor)
contrasts(CARS.factor)

  #### Create TONS variable
Tons_box <- ggplot(totacts_wCasualties_DR_ER, aes(y=TONS)) + geom_boxplot()

med <- ggplot_build(Tons_box)$data[[1]]$middle

TONS.factor <- totacts_wCasualties_DR_ER$TONS
TONS.factor[which(totacts_wCasualties_DR_ER$TONS<med)]<-'Low tonnage'
TONS.factor[which(totacts_wCasualties_DR_ER$TONS>=med)]<-'high tonnage'
TONS.factor <- factor(TONS.factor)
contrasts(TONS.factor)

### Plot
  #### Variables of interest: day, clear, HwyRail, Derail, Pass_or_Comm, TRNSPD.factor, CARS.factor, TONS.factor

  #### Highway Rail Plots
  # Interaction of highway rail and day
interaction.plot(HwyRail, day, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and clear
interaction.plot(HwyRail, clear, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and Pass_or_Comm
interaction.plot(HwyRail, Pass_or_Comm, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and TRNSPD.factor
interaction.plot(HwyRail, TRNSPD.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and CARS.factor
interaction.plot(HwyRail, CARS.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and TONS.factor
interaction.plot(HwyRail, TONS.factor, totacts_wCasualties_DR_ER$Casualty)



#### Derailment Plots
  # Interaction of highway rail and day
interaction.plot(Derail, day, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and clear
interaction.plot(Derail, clear, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and Pass_or_Comm
interaction.plot(Derail, Pass_or_Comm, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and TRNSPD.factor
interaction.plot(Derail, TRNSPD.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and CARS.factor
interaction.plot(Derail, CARS.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and TONS.factor
interaction.plot(Derail, TONS.factor, totacts_wCasualties_DR_ER$Casualty)


#### Passenger or Commuter
  # Interaction of Passenger or Commuter and day
interaction.plot(Pass_or_Comm, day, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of Passenger or Commuter and clear
interaction.plot(Pass_or_Comm, clear, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of Passenger or Commuter and TRNSPD.factor
interaction.plot(Pass_or_Comm, TRNSPD.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of Passenger or Commuter and CARS.factor
interaction.plot(Pass_or_Comm, CARS.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of Passenger or Commuter and TONS.factor
interaction.plot(Pass_or_Comm, TONS.factor, totacts_wCasualties_DR_ER$Casualty)

print("Looking at highway-rail interactions, it seems like non-highway rail accidents are worse on average which makes sense given the previous observation that derailment accidents in particular are worse per accident on average.")

print("Looking at derailment interactions, low tonnage accidents are worse on average which is probably because passengar cars weigh less than non-passenger cars on average. This interaction doesn't tell us much but we may still choose to model it. The involvement of one or more hazard cars increases average casualties so this might be good to include. High train speed clearly has a strong interaction with derailment type accidents. Passenger or commuter train involvement increase severity. Most severe accidents happen during clear weather. Finally, daytime accidents are more severe on average.")

print("Looking at passenger or commuter train type accidents, lower tonnage is associated with more severe accident damage. When hazard cars are present, accident severity seems to increase. There is an interaction between passenger or commuter and train speed. Accident severity seems to increase with bad weather.")


###### Most important plots pasted for convenience:
  # 3rd PC Loadings
barplot(totacts_wCasualties.pca$loadings[,3], main='PC3 Loadings')

  # Interaction of highway rail and TONS.factor
interaction.plot(Derail, TONS.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and CARS.factor
interaction.plot(Derail, CARS.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and TRNSPD.factor
interaction.plot(Derail, TRNSPD.factor, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of highway rail and Pass_or_Comm
interaction.plot(Derail, Pass_or_Comm, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of Passenger or Commuter and day
interaction.plot(Pass_or_Comm, day, totacts_wCasualties_DR_ER$Casualty)

  # Interaction of Passenger or Commuter and CARS.factor
interaction.plot(Pass_or_Comm, CARS.factor, totacts_wCasualties_DR_ER$Casualty)

ggarrange(plt_1, plt_2, plt_3, plt_4, plt_5, plt_6, nrow = 2, ncol = 3)
ggarrange(plt_7, plt_8, plt_9, plt_10, plt_11, plt_12, nrow = 2, ncol = 3)

############### Playing Around

```

# Hypothesis formation:
Hypothesis 1:
$H_0:$ There is no relationship or an inverse relationship between derailment type accidents with passenger or commuter cars and accident severity relative to other types of accidents with other types of cars.
$H_1:$ There is a positive relationship between derailment type accidents with passenger or commuter cars and accident severity relative to other types of accidents with other types of cars.

Hypothesis 2:
$H_0:$ There is no interaction between derailment type accidents involving passenger or commuter cars and the number of hazard cars involved and train speed affecting accident severity.
$H_1:$ There is an interaction between derailment type accidents involving passenger or commuter cars and the number of hazard cars involved and train speed affecting accident severity.

